Fifteen years of decentralization attempts

### 2014-03-24 paper.tex
Fixed typos

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	In the emerging world of ubiquitous internet access, void of any privacy, there are a number of anonymous communication overlay networks whose aim it is to provide privacy to its users. Due to the nature of the internet, there is major difficulty in getting these networks to become both decentralized and anonymous. This technical report lists reasons for having anonymous networks, discerns the problems in achieving decentralization in anonymous networks and sums up the biggest competitors in the field and their current status. To do so, we use one exemplary network, the Tor network. We explain how Tor works, what vulnerabilities this network currently has, and possible attacks that could be used to violate privacy and anonymity. The Tor network is used as a key comparison network in the main part of the report: a tabular overview of the major anonymous networking technologies in use today.
	%There are quite a few distributed anonymous communication networks such as Tor, yet when it comes to decentralized anonymous systems the numbers are rather small and certain problems arise. In this technical report we explain how Tor works and which attacks are possible and which vulnerabilities currently exist in Tor. After that, we talk about why would like to decentralize peer-to-peer networks, what the difficulties are with decentralizing networks and show which projects have made an attempt to decentralize or are already (partially) decentralized. Finally, we provide a table in where we compare these projects against each other, which will also be public available on the Github page mention in this report.
\end{abstract}

@@

	Although much research has been done on anonymous internet communication, only few systems have been actually implemented and are being used in practice. One of the most important factors that impact anonymity in a communication system is the number of users. A sufficiently large numbers of users are required for the system to make guarantees about its ability to conceal and hide users' communication. This makes it difficult to introduce a new anonymous internet system, due to the initial lack of users. Another important factor that impacts the anonymity of the system is decentralization. Whenever a network has centralized components and a malicious third party gains control over these components, the functionality and perhaps even the anonymity of the system is compromised.

@@

	\begin{figure*}[!t]
		\centering
		\includegraphics[width=0.8\textwidth]{graphics/tor.pdf}
		\caption{The components of the Tor network. After downloading the node list from the Directory Server, the user creates a circuit through a guard node, a relay node and an exit node, and uses it to communicate (anonymously) with the internet.}
		\label{fig:tor_layout}
	\end{figure*}

@@

		\begin{itemize}
			\item Tor's congestion control does not work well. The network has some problems handling bulk transfers, such as downloading large files or streaming high-quality videos. The congestion control could be improved by using an unreliable protocol for links between Tor relays. Goldberg et al. \cite{alsabah2013pctcp} have proposed PCTCP which could improve the response time of the Tor network by 60\% and the download time of files by 30\%.

@@

			\item The Tor network doesn't have the capacity to handle all the users that want privacy on the Internet. According to the Tor Metrics project \cite{tormetricsprojectwebsite}, it takes about 6 seconds to download 1 MiB of data. By increasing the amount of Onion Routing in the Tor network, the capacity is increased. Incentives such as LIRA \cite{jansen13lira} could make more users run an Onion Router, thus increasing the capacity of the network and making it faster.

@@

	\subsection{Low-Resource Routing attacks}
		Another possible attack on the anonymity of tor is the so called \emph{low-resource routing attack} \cite{bauer2007low}. With this attack, it is possible for an adversary to perform an end-to-end traffic analysis with minimal resources, thus compromising the anonymity Tor provides. The idea of this attack is that a malicious Onion router can lie about its bandwidth, thus advertising a much higher bandwidth than it actually has. Because of the Tor path selection algorithm that prefers high-bandwidth nodes, the chance that a malicious entry and exit node are chosen, is high.

		Once a malicious entry and exit node have been chosen, an analysis of the traffic can be done to link Onion proxies with the web servers they communicate with. Experimental research in a test setting has shown that with an total of 66 non-malicious and 6 malicious nodes, is possible to compromise 46\% of the built circuits. At the request of the Tor community, this attack hasn't been tested on the live Tor network. There are some extensions and improvements to this attack. It is possible to perform a denial of service attack on well used entry nodes, forcing Tor clients to choose a new one. This improves the chance that a Tor client chooses a malicious entry node.

@@

		Using this, an accomplice exit router can, in cooperation with the entry node, discover the sender and receiver's relationship. This attack can also be used as a denial of service attack. According to Pries et al., defending against this attack is quite challenging to solve and requires further research.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	In the emerging world of ubiquitous internet access, void of any privacy, there are a number of anonymous communication overlay networks whose aim it is to provide privacy to its users. Due to the nature of the internet, there is major difficulty in getting these networks to become both decentralized and anonymous. This technical report lists reasons for having anonymous networks, discerns the problems in achieving decentralization in anonymous networks and sums up the biggest competitors in the field and their current status. To do so, we use one exemplary network, the Tor network. We explain how Tor works, what vulnerabilities this network currently has and possible attacks that could be used to violate privacy and anonymity. The Tor network is used as a key comparison network in the main part of the report: a tabular overview of the major anonymous networking technologies in use today.
	%There are quite a few distributed anonymous communication networks such as Tor, yet when it comes to decentralized anonymous systems the numbers are rather small and certain problems arise. In this technical report we explain how Tor works and which attacks are possible and which vulnerabilities currently exist in Tor. After that, we talk about why would like to decentralize peer-to-peer networks, what the difficulties are with decentralizing networks and show which projects have made an attempt to decentralize or are already (partially) decentralized. Finally, we provide a table in where we compare these projects against each other, which will also be public available on the Github page mention in this report.
\end{abstract}

@@

	Although much research has been done on anonymous internet communication, only few systems have been actually implemented and are being used in practice. One of the most important factors that impacts anonymity in a communication system is the number of users. A sufficiently large numbers of users are required for the system to make guarantees about its ability to conceal and hide users' communication. This makes it difficult to introduce a new anonymous internet system, due to the initial lack of users. Another important factor that impacts the anonymity of the system is decentralization. Whenever a network has centralized components and a malicious third party gains control over these components, the functionality and perhaps even the anonymity of the system is compromised.

@@

	\begin{figure*}[!t]
		\centering
		\includegraphics[width=0.8\textwidth]{graphics/tor.pdf}
		\caption{The components of the Tor network. After downloading the node list from the Directory Server, the user creates a circuit through a guard node, a relay node and an exit node. This circuit is used to communicate (anonymously) with the internet.}
		\label{fig:tor_layout}
	\end{figure*}

@@

		\begin{itemize}
			\item Tor's congestion control does not work well. The network has some problems handling bulk transfers, such as downloading large files or streaming high-quality videos. The congestion control could be improved by using an unreliable protocol such as UDP for links between Tor relays. Goldberg et al. \cite{alsabah2013pctcp} have proposed PCTCP which could improve the response time of the Tor network by 60\% and the download time of files by 30\%.

@@

			\item The Tor network doesn't have the capacity to handle all the users that want privacy on the internet. According to the Tor Metrics project \cite{tormetricsprojectwebsite}, it takes about 6 seconds to download 1 MiB of data. By increasing the amount of Onion Routing in the Tor network, the capacity is increased. Incentives such as LIRA \cite{jansen13lira} could make more users run an Onion Router, thus increasing the capacity of the network and making it faster.

@@

	\subsection{Low-Resource Routing attacks}
		Another possible attack on the anonymity of Tor is the so called \emph{low-resource routing attack} \cite{bauer2007low}. With this attack, it is possible for an adversary to perform an end-to-end traffic analysis with minimal resources, thus compromising the anonymity Tor provides. The idea of this attack is that a malicious Onion router can lie about its bandwidth, thus advertising a much higher bandwidth than it actually has. Because of the Tor path selection algorithm that prefers high-bandwidth nodes, the chance that a malicious entry and exit node are chosen, is high.

		Once a malicious entry and exit node have been chosen, an analysis of the traffic can be done to link Onion proxies with the web servers they communicate with. Experimental research in a test setting has shown that with a total of 66 non-malicious and 6 malicious nodes, is possible to compromise 46\% of the built circuits. At the request of the Tor community, this attack hasn't been tested on the live Tor network. There are some extensions and improvements to this attack. It is possible to perform a denial of service attack on well used entry nodes, forcing Tor clients to choose a new one. This improves the chance that a Tor client chooses a malicious entry node.

@@

		Using this, an accomplice exit router can, in cooperation with the entry node, discover the sender and receiver's relationship. This attack can also be used as a denial of service attack. According to Pries et al., defending against this attack is quite challenging and requires further research.
